---
title: "Variants - CaseComparison"
author: "Fouzia Farooq"
date: "8/5/2022"
output: pdf_document
---

Emily asked to generate a figure for the case-comparison paper showing how much data we have from our studies that were included in the CC paper - most only overlap Pre-alpha period.  BMJ reviewers had a comment and wanted us to do the analysis. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(ggplot2)
library(RColorBrewer)
library(readxl)
library(xlsx)
library(patchwork)
library(haven)
library(meta)
# library(metafor)
# library(esc)
```

NOTES: json file was converted to csv in python (deepnote.com), then copied the code into BBtext and ran at the command line. 
Then ran the following code. 
Dates are fractions of a year after "year." We a creating a seperate dataframe into a csv that we will bring in R and use lubridate to read the date properly. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


OPENSOURCE
```{r}
frequency <- read.csv('data/Nextstrain_opensource_frequenices.csv')
covid_opensource <- read.csv('data/nextstrain_ncov_open_global_all-time_metadata_20220517.csv')
pivots <- read.csv('data/Nextstrain_opensource_pivotdates.csv', header = FALSE)
study_cases <- read.csv('data/variant_study_case_control_COVID_data_ff_20220805.csv') # case-comparison country study dates

# Study files
# compiled_ipd_data_old <- read_dta('data/Compiled_IPD_Data_220217.dta') # old ipd data file
# compiled_ipd_data_new <- read_dta('data/Compiled_IPD_Data_220620.dta') # updated with June 2022 data
# will take all the patient IDs from the new file and then add in the missing rows that are in the old file to this file. 
# agd_wide <- read_dta('data/Compiled_AGD_WIDE_220408.dta') # Has events, country, but no dates.
# agd_long <- read_dta('data/Compiled_AGD_220408.dta') # Has events, country, but no dates.
country_id <- read.csv('data/country_id_ipd.csv') # FF created this file
# country_id_agd <- read.csv('data/country_id_agd.csv')
```


```{r}
opensource.df <- covid_opensource %>% dplyr::select(strain, date, country, clade_membership)

opensource.df2 <- left_join(frequency, covid_opensource)

pivot_df <- data.frame(week = 0:(length(pivots$V1)-1), pivot = pivots$V1)
pivot_df <- pivot_df %>% dplyr::mutate(date = as.Date(date_decimal(pivot))) #converting the dates in decimal in nextstrain dataset to readable dates.

df <- left_join(opensource.df2, pivot_df, by="week")
```


```{r}
df2 <- df  %>% dplyr::mutate(variant = case_when(clade_membership %in% 
                                                   c("19A", "19B", "20A", "20B", "20C", "20D", "20F", "20G", "20E (EU1)") ~ "Pre_alpha",
                                                 clade_membership %in% 
                                                   c("20I (Alpha, V1)") ~ "Alpha",
                                                 clade_membership %in%
                                                   c("20H (Beta, V2)") ~ "Beta", 
                                                 clade_membership %in%
                                                   c("21B (Kappa)") ~ "Kappa", 
                                                 clade_membership %in%
                                                   c("21D (Eta)") ~ "Eta",
                                                 clade_membership %in% 
                                                   c("21E (Theta)") ~ "Theta",
                                                 clade_membership %in% 
                                                   c("20J (Gamma, V3)") ~ "Gamma",
                                                 clade_membership %in% 
                                                   c("21A (Delta)", "21I (Delta)", "21J (Delta)") ~ "Delta",
                                                 clade_membership %in% 
                                                   c("21C (Epsilon)") ~ "Epsilon",
                                                 clade_membership %in% 
                                                   c("21F (Iota)") ~ "Iota",
                                                 clade_membership %in% 
                                                   c("21G (Lambda)") ~ "Lambda",
                                                 clade_membership %in% 
                                                   c("21H (Mu)") ~ "Mu",
                                                 clade_membership %in% 
                                                   c("21K (Omicron)", "21L (Omicron)", "21M (Omicron)", "22C (Omicron)") ~ "Omicron"))

# table(df2$clade_membership, df2$variant, useNA = "always") # These all have a general variant name now. 

```

NORMALIZE FREQ BY COUNTRY
```{r}
df3 <- df2 %>% group_by(country, week, variant) %>% summarize(freq = sum(frequency))
check.df <- df3 %>% filter(country=="Africa") # Nextstrain does not have data on DRC, Africa also can't be listed. 
df4 <- df3 %>% group_by(country, week) %>% 
  dplyr::mutate(nor = (freq/sum(freq))) %>%
  dplyr::filter(country %in% c("Kenya", "Uganda", "South Africa", "Ghana", "Nigeria", 
                               "Spain", "Hong Kong", "Italy", "USA", "Sweden", "Turkey")) # List does not include DRC or Burkina Faso b/c opensource data is not available thorugh Nextstrain.  This list doesn't include all the african countries or the WHO countries. 

```


```{r}
df5 <- left_join(df4, pivot_df, by="week")

df6 <- df5 %>% dplyr::mutate(date_ym = format(date, format= "%y-%m"))
df7 <- df6 %>% group_by(country, date_ym) %>% dplyr::mutate(max_date = max(date))
df8 <- df7 %>% dplyr::filter(date == max_date)
# sort(colnames(compiled_ipd_data))
```


Joining RFA dates only - NO NEED TO RUN FOR RFA ONLY (may need for CC):
```{r}
study_cases$date <- mdy(study_cases$date)
cases <- study_cases %>% dplyr::mutate(date_ym = format(date,format="%y-%m"))
class(study_cases$date)

variants_cases <- left_join(df8, cases, by=c("country", "date_ym"))
```

New map with overlayed CC study dates  dates: 06/27/2022 - latest
```{r}
studies_df <- variants_cases %>% dplyr::filter(Events>=1) %>% 
  group_by(country, author_year) %>% # had "site_name" instead of "author_year" before. Also had "study_type" here. 
  summarize(min_month = min(pandemic_month),
            max_month = max(pandemic_month)) %>%
  group_by(country) %>%
  mutate(country_study_number = 1:n(),
         country_total_studies = n())
# see https://stackoverflow.com/questions/36893957/dplyr-generate-row-number-row-position-in-group-by

gap_betw_studies <- 0.15

# Splitting up into 3 pages.
page1_df <- studies_df %>% dplyr::filter(country %in% c( "Ghana", "Hong Kong", "Italy", "Kenya"))
page2_df <- studies_df %>% dplyr::filter(country %in% c("Nigeria", "South Africa", "Spain", "Sweden"))
page3_df <- studies_df %>% dplyr::filter(country %in% c("Turkey", "Uganda", "USA"))

page_list_df <- list(page1_df, page2_df, page3_df)

page1_df8 <- df8 %>% dplyr::filter(country %in% c( "Ghana", "Hong Kong", "Italy", "Kenya"))
page2_df8 <- df8 %>% dplyr::filter(country %in% c("Nigeria", "South Africa", "Spain", "Sweden"))
page3_df8 <- df8 %>% dplyr::filter(country %in% c("Turkey", "Uganda", "USA"))

page_list_df8 <- list(page1_df8, page2_df8, page3_df8)

for(page_num in 1:length(page_list_df)) { # will take one df off the list each time. 
  
  ggplot() +
    geom_bar(data = page_list_df8[[page_num]], stat="identity", #pull out either part1,2, or 3. had "df8" here before. 
             mapping = aes(x = date_ym, fill = variant, y=nor)) +
    scale_fill_manual(values = c("Alpha" = "deepskyblue2", "Pre_alpha" = "lightsteelblue1", "Beta" = "darkseagreen1",
                                 "Delta" = "gold", "Epsilon" = "burlywood", "Eta" = "cadetblue2", "Gamma" = "darkgoldenrod1",
                                 "Iota" = "cyan3", "Lambda" = "gray70", "Mu" = "peachpuff1", "Omicron" = "plum2",
                                 "Theta" = "tomato")) + 
    theme_light() + 
    theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size = 6),
          axis.text.y = element_text(size = 6)) + 
    xlab("Year-Month") + 
    ylab("Relative Frequency") +
    geom_rect(data = page_list_df[[page_num]], # had studies_df here before.
              mapping=aes(xmin=min_month-0.5,
                          xmax=max_month+0.5,
                          ymin= 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + gap_betw_studies*(country_study_number - 1) - 0.005,
                          ymax= 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + gap_betw_studies*(country_study_number - 1) + 0.005)) +
    geom_text(data = page_list_df[[page_num]], # had studies_df here before.
              mapping=aes(x = 0.5*(min_month + max_month),
                          y = 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) +
                            gap_betw_studies*(country_study_number - 1) + 0.05,
                          label = author_year),
              size = 2) +
              # label = ifelse(is.na(author_year))),
              #study_type,
              # paste(author_year, study_type))),
              
    guides(fill = guide_legend(override.aes = list(labels = c("A", "B")))) +
    theme(aspect.ratio = 0.3) + # makes all the facets the same size (b/c page 3 has only 3 whereas pages 1,2 have 4 facets)
    facet_wrap(~country, nrow=4, ncol = 1, scale = "free_x", drop=FALSE) #drop
  
  ggsave(filename = paste0('data_out/covid_variant_CC-countries_20220809_', page_num, '.pdf'), width = 8.5, height = 11, units = "in") # DON'T NEED TO WRITE THIS OUT EVERYTIME.
}
# write.csv(df8, 'data/nextstrain_frequency_data_written_ff_20220610.csv', row.names = FALSE)
```


