---
title: "Variants - CaseComparison"
author: "Fouzia Farooq"
date: "8/5/2022"
output: pdf_document
---

Emily asked to generate a figure for the case-comparison paper showing how much data we have from our studies that were included in the CC paper - most only overlap Pre-alpha period.  BMJ reviewers had a comment and wanted us to do the analysis. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(ggplot2)
library(RColorBrewer)
library(readxl)
library(xlsx)
library(patchwork)
library(haven)
library(meta)
library(R.utils)
# library(metafor)
# library(esc)
```

NOTES: json file was converted to csv in python (deepnote.com), then copied the code into BBtext and ran at the command line. 
Then ran the following code. 
Dates are fractions of a year after "year." We a creating a seperate dataframe into a csv that we will bring in R and use lubridate to read the date properly. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


OPENSOURCE
```{r}
# frequency <- read.csv('data/Nextstrain_opensource_frequenices.csv')
# covid_opensource <- read.csv('data/nextstrain_ncov_open_global_all-time_metadata_20220517.csv')
# pivots <- read.csv('data/Nextstrain_opensource_pivotdates.csv', header = FALSE)
# study_cases <- read.csv('data/variant_study_case_control_COVID_data_ff_20220805.csv') # case-comparison country study dates
# 
# # Study files
# # compiled_ipd_data_old <- read_dta('data/Compiled_IPD_Data_220217.dta') # old ipd data file
# # compiled_ipd_data_new <- read_dta('data/Compiled_IPD_Data_220620.dta') # updated with June 2022 data
# # will take all the patient IDs from the new file and then add in the missing rows that are in the old file to this file. 
# # agd_wide <- read_dta('data/Compiled_AGD_WIDE_220408.dta') # Has events, country, but no dates.
# # agd_long <- read_dta('data/Compiled_AGD_220408.dta') # Has events, country, but no dates.
# country_id <- read.csv('data/country_id_ipd.csv') # FF created this file
# # country_id_agd <- read.csv('data/country_id_agd.csv')
```


GISAID Data
```{r}
# Control parameters
data_source <- 'gisaid' # 'open'
redownload <- FALSE # switch to TRUE to get fresh data

if (redownload) {
  # Download data and gunzip
  download.file(url = paste0('https://data.nextstrain.org/files/workflows/forecasts-ncov/',
                             data_source, '/nextstrain_clades/global.tsv.gz'),
                destfile = 'data/global.tsv.gz')
  
  gunzip('data/global.tsv.gz', remove = FALSE, overwrite = TRUE)
}

gisaid_df <- read_delim('data/global_20240319.tsv',
                 delim = '\t') %>%
  rename(country = location)

# Extract 2-digit year-month (20-12)
gisaid_df <- gisaid_df %>%
  mutate(date_ym = substr(date, 3, 7))

# table(gisaid_df$country)
```

Read in the data file
```{r}
CC_data <- read.xlsx('data/Total sample size_11-2024.xlsx', sheetName = 'Sheet1')
```


***GISAID DATA***
Nomenclature for new variants still called Omicron by WHO: 
https://covid.cdc.gov/covid-data-tracker/#variant-summary
```{r}
# The clade categories are updated Dec 13, 2023.
gisaid_df <- gisaid_df %>% 
  mutate(variant = case_when(clade %in% 
                               c("19A", "19B", "20A", "20B", "20C", "20D", "20F", "20G", "20E") ~ "Pre_alpha",
                             clade %in% 
                               c("20I") ~ "Alpha",
                             clade %in%
                               c("20H") ~ "Beta", 
                             clade %in%
                               c("21B") ~ "Kappa", 
                             clade %in%
                               c("21D") ~ "Eta",
                             clade %in% 
                               c("21E") ~ "Theta",
                             clade %in% 
                               c("20J", "20J") ~ "Gamma",
                             clade %in% 
                               c("21A", "21I", "21J") ~ "Delta",
                             clade %in% 
                               c("21C") ~ "Epsilon",
                             clade %in% 
                               c("21F") ~ "Iota",
                             clade %in% 
                               c("21G") ~ "Lambda",
                             clade %in% 
                               c("21H") ~ "Mu",
                             clade %in% 
                               c("21K", # BA.1
                                 "21L", # BA.2
                                 "21M", # Omicron
                                 "22A", # BA.4
                                 "22B", # BA.5
                                 "22C", # BA.2.12.1
                                 "22D", # BA.2.75
                                 "22E", # BQ.1
                                 "23C", # CH.1.1
                                 "23I", # BA.2.86 - highly mutated vs. Omicron but emerged late 2023
                                 
                                 # These below were set to XBB but I added them to Omicron. 
                                 "22F", # XBB
                                 "23A", # XBB.1.5
                                 "23B", # XBB.1.16
                                 "23D", # XBB.1.9
                                 "23E", # XBB.2.3
                                 "23F", # EG.5.1
                                 "23G", # XBB.1.5.70
                                 "23H", # HK.3
                                 # Added 2024-11-26:
                                 "24A",
                                 "24B", 
                                 "24C",
                                 "24D",
                                 "24E",
                                 "24F",
                                 "24G",
                                 "24I")
                             ~ "Omicron",
                             clade %in%
                               c("recombinant") ~ "recombinant"))

```

# Review missing clade data in GISAID
```{r}
temp.df <- gisaid_df %>% filter(is.na(variant)) %>% filter(country %in% CC_data$Country)
range(as_date(temp.df$date))
```


# New map with overlaid CC study dates and CC study dates: 11-26-2024 - latest
```{r fig.height=12, fig.width=8}

studies_df <- CC_data %>% 
  mutate(startdate_ym = substr(start_date, 3, 7),
         enddate_ym = substr(end_date, 3, 7)) %>%
  group_by(country) %>%
  mutate(country_study_number = 1:n(),
         country_total_studies = n())
# see https://stackoverflow.com/questions/36893957/dplyr-generate-row-number-row-position-in-group-by


# Compute monthly totals by country and variant
df_monthly <- gisaid_df %>%
  mutate(date_ym = substr(date, 3, 7)) %>%
  group_by(date_ym, country, variant) %>%
  summarize(count = sum(sequences, na.rm = TRUE))

# Compute relative ("normalized") variant frequencies per month, location
df_monthly <- df_monthly %>%
  group_by(date_ym, country) %>%
  mutate(nor = round(count/sum(count, na.rm = TRUE), digits = 3))

max(studies_df$enddate_ym) # 23-07

# Truncate strain data after 2023
df_monthly <- df_monthly %>%
  filter(date_ym < "24-01") %>%
  filter(country %in% studies_df$country)
```

```{r}
gap_betw_studies <- 0.17
studies_y_nudge <- -0.04
all_dates <- names(table(df_monthly$date_ym))

ym_lookup <- data.frame(date_ym = all_dates, 
                        pandemic_month = 1:length(all_dates))

studies_df <- left_join(studies_df, ym_lookup, by = c("startdate_ym" = "date_ym")) %>%
  rename(min_month = pandemic_month)

studies_df <- left_join(studies_df, ym_lookup, by = c("enddate_ym" = "date_ym")) %>%
  rename(max_month = pandemic_month)

ggplot() +
  geom_bar(data = df_monthly, 
           stat="identity",
           width = 0.8,
           mapping = aes(x = date_ym, fill = variant, y=nor)) +
  scale_fill_manual(values = c("Alpha" = "deepskyblue2", "Pre_alpha" = "lightsteelblue1", "Beta" = "darkseagreen1",
                               "Delta" = "gold", "Epsilon" = "burlywood", "Eta" = "cadetblue2", "Gamma" = "darkgoldenrod1",
                               "Iota" = "cyan3", "Lambda" = "gray70", "Mu" = "peachpuff1", "Omicron" = "plum2",
                               "Theta" = "tomato")) + 
  theme_light() + 
  theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size = 5),
        axis.text.y = element_text(size = 5.5),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) + 
  xlab("Year-Month") + 
  ylab("Relative Frequency") +
  geom_rect(data = studies_df,
            mapping=aes(xmin=min_month-0.5,
                        xmax=max_month+0.5,
                        ymin= 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + 
                          gap_betw_studies*(country_study_number - 1) - 0.005 + studies_y_nudge,
                        ymax= 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + 
                          gap_betw_studies*(country_study_number - 1) + 0.005 + studies_y_nudge)) +
  geom_text(data = studies_df,
            mapping=aes(x = 0.5*(min_month + max_month),
                        y = 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + 
                          gap_betw_studies*(country_study_number - 1) + 0.06 + studies_y_nudge,
                        label = author_year),
            size = 1.6) +
  guides(fill = guide_legend(override.aes = list(labels = c("A", "B")))) +
  xlim(all_dates) + 
  facet_wrap(~country, ncol = 1, scale = "free_x")

today_date <- format(Sys.time(), "%Y%m%d")
ggsave(filename = paste0('plots/variants_gisaid/covid_variant_country_',today_date, '.pdf'), width = 6, height = 30, units = "in", 
       limitsize = FALSE) 

#TODO Reroder alpha beta theta legend.
```





