---
title: "COVID-19_VariantStudy_v1"
author: "Fouzia Farooq"
date: "5/11/2022"
output: pdf_document
---

```{r}
library(dplyr)
library(readr)
library(lubridate)
library(ggplot2)
library(RColorBrewer)
library(readxl)
```

NOTES: json file was converted to csv in python (deepnote.com), then copied the code into BBtext and ran at the command line. 
Then ran the following code. 
Dates are fractions of a year after "year." We a creating a seperate dataframe into a csv that we will bring in R and use lubridate to read the date properly. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- ```{r} -->
<!-- usa <- read_delim('data/nextstrain_ncov_open_global_all-time_metadata.tsv', col_types = cols(date = col_date("%m/%d/%y"), -->
<!--                                                                                              division = "f")) -->
<!-- covid.data <- read_delim('data/nextstrain_ncov_gisaid_global_all-time_metadata.tsv', col_types = cols(date = col_date("%Y-%m-%d"))) -->
<!-- ``` -->


<!-- GISAID:  -->
<!-- ```{r} -->
<!-- frequency <- read.csv('data/Nextstrain_frequenices.csv') -->
<!-- gisaid <- read.csv('data/metadata_gisaid_20220516.csv') -->
<!-- pivots <- read.csv('data/Nextstrain_pivotdates.csv', header = FALSE) -->

<!-- gisaid.df <- gisaid %>% dplyr::select(strain, date, country, clade_membership) -->

<!-- gisaid.df2 <- left_join(frequency, gisaid.df) -->

<!-- clade.freq.week <- gisaid.df2 %>% group_by(country, week, clade_membership) %>% summarize(freq = sum(frequency)) -->

<!-- pivot_df <- data.frame(week = 0:(length(pivots$V1)-1), pivot = pivots$V1) -->
<!-- pivot_df <- pivot_df %>% dplyr::mutate(date = as.Date(date_decimal(pivot))) -->

<!-- gisaid_clade <- left_join(clade.freq.week, pivot_df) -->
<!-- ``` -->

<!-- Study cases data:  -->
<!-- ```{r} -->
<!--   cases <- study_cases %>% dplyr::mutate(year = as.integer(substring(Strata, 10, 13)), -->
<!--                       month = as.integer(substring(Strata, 14,15))) -->
<!--   cases <- cases %>% dplyr::mutate(pandemic_month = month + (year-2020)*12) -->
<!-- ``` -->

OPENSOURCE
```{r}
frequency <- read.csv('data/Nextstrain_opensource_frequenices.csv')
covid_opensource <- read.csv('data/nextstrain_ncov_open_global_all-time_metadata_20220517.csv')
pivots <- read.csv('data/Nextstrain_opensource_pivotdates.csv', header = FALSE)
# study_cases <- read.csv('data/casesbymonth_220329_ff.csv', header = TRUE, na.strings = "") # old file
study_cases <- read_xlsx('data/variant_study_case_control_COVID_data_ff_20220524.xlsx') # case-comparison country study dates
```

```{r}
opensource.df <- covid_opensource %>% dplyr::select(strain, date, country, clade_membership)

opensource.df2 <- left_join(frequency, covid_opensource)


pivot_df <- data.frame(week = 0:(length(pivots$V1)-1), pivot = pivots$V1)
pivot_df <- pivot_df %>% dplyr::mutate(date = as.Date(date_decimal(pivot))) #converting the dates in decimal in nextstrain dataset to readable dates.

df <- left_join(opensource.df2, pivot_df, by="week")
```


```{r}
df2 <- df  %>% dplyr::mutate(variant = case_when(clade_membership %in% 
                                                   c("19A", "19B", "20A", "20B", "20C", "20D", "20F", "20G", "20E (EU1)") ~ "Pre_alpha",
                                                 clade_membership %in% 
                                                   c("20I (Alpha, V1)") ~ "Alpha",
                                                 clade_membership %in%
                                                   c("20H (Beta, V2)") ~ "Beta", 
                                                 clade_membership %in%
                                                   c("21B (Kappa)") ~ "Kappa", 
                                                 clade_membership %in%
                                                   c("21D (Eta)") ~ "Eta",
                                                 clade_membership %in% 
                                                   c("21E (Theta)") ~ "Theta",
                                                 clade_membership %in% 
                                                   c("20J (Gamma, V3)") ~ "Gamma",
                                                 clade_membership %in% 
                                                   c("21A (Delta)", "21I (Delta)", "21J (Delta)") ~ "Delta",
                                                 clade_membership %in% 
                                                   c("21C (Epsilon)") ~ "Epsilon",
                                                 clade_membership %in% 
                                                   c("21F (Iota)") ~ "Iota",
                                                 clade_membership %in% 
                                                   c("21G (Lambda)") ~ "Lambda",
                                                 clade_membership %in% 
                                                   c("21H (Mu)") ~ "Mu",
                                                 clade_membership %in% 
                                                   c("21K (Omicron)", "21L (Omicron)", "21M (Omicron)", "22C (Omicron)") ~ "Omicron"))

table(df2$clade_membership, df2$variant, useNA = "always") # These all have a general variant name now. 

```

NORMALIZE FREQ BY COUNTRY
```{r}
df3 <- df2 %>% group_by(country, week, variant) %>% summarize(freq = sum(frequency))
df4 <- df3 %>% group_by(country, week) %>% 
  dplyr::mutate(nor = (freq/sum(freq))) %>%
  dplyr::filter(country %in% c("Kenya", "Uganda", "Spain", "Hong Kong", "Italy",
                               "South Africa", "USA", "Sweden", "Turkey", "China", "United Kingdom", "Chile", "Canada", 
                                "Mexico", "Colombia", "USA", "India", "Africa", "Kenya")) # List does not include DRC or Burkina Faso b/c opensource data is not available.  thorugh Nextstrain.  This list doesn't include all the african countries or the WHO countries. 

```


<!-- ```{r} -->
<!-- # this is global data - it matches the nextstrain viz.  -->
<!-- clade.freq.week <- opensource.df2 %>% group_by(week, clade_membership) %>% summarize(freq = sum(frequency)) -->
<!-- opensource.df3 <- left_join(clade.freq.week, pivot_df, by="week") -->
<!-- ``` -->




```{r}
df5 <- left_join(df4, pivot_df, by="week")

df6 <- df5 %>% dplyr::mutate(date_ym = format(date, format= "%y-%m"))
df7 <- df6 %>% group_by(country, date_ym) %>% dplyr::mutate(max_date = max(date))
df8 <- df7 %>% dplyr::filter(date == max_date)
```

<!-- Joining on the study cases data -->
<!-- ```{r} -->
<!-- cases2 <- cases %>% dplyr::select(Events, country=Country, year, month, pandemic_month) %>%  -->
<!--   dplyr::mutate(date_ym = paste0(as.character(year-2000), "-", sprintf("%02d", month))) -->

<!-- variants_cases <- left_join(df8, cases2, by=c("country", "date_ym")) -->
<!-- ``` -->

Joining the case-comparison dates:
```{r}
study_cases$date <- mdy(study_cases$date)
cases <- study_cases %>% dplyr::mutate(date_ym = format(date,format="%y-%m"))
class(study_cases$date)

variants_cases <- left_join(df8, cases, by=c("country", "date_ym"))
```

Creates the older map: 
```{r}
fill_colors <- factor(1,2)
palette1 <- c("#999999", "000000")
variants_cases %>% ggplot() +
  geom_bar(stat="identity",
           mapping = aes(x = date_ym, fill = variant, y=nor)) +
  scale_fill_manual(values = c("Alpha" = "deepskyblue2", "Pre_alpha" = "gray70", "Beta" = "darkseagreen1",
                               "Delta" = "gold", "Epsilon" = "burlywood", "Eta" = "cadetblue2", "Gamma" = "darkgoldenrod1",
                               "Iota" = "cyan3", "Lambda" = "lightsteelblue1", "Mu" = "peachpuff1", "Omicron" = "plum2",
                               "Theta" = "tomato", "present" = rgb(0,0,0, alpha=0.5), "absent" = rgb(0,0,0,alpha = 0))) + 
  theme_light() + 
  theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size = 6),
        axis.text.y = element_text(size = 6)) + 
  geom_rect(mapping=aes(xmin=pandemic_month-0.5,
                        xmax=pandemic_month+0.5,
                        ymin=0.48,
                        ymax=0.52,
                        fill = ifelse(Events>0, "present", "absent"))) +
  guides(fill = guide_legend(override.aes = list(labels = c("A", "B")))) +
  facet_wrap(~country, nrow = 30, ncol = 1, scale = "free_x")
   
ggsave(filename = 'data_out/covid_variant_country_20220524.pdf', width = 6, height = 25, units = "in")
  
```

New map with overlayed CC study dates and RFA study dates: 
```{r}

studies_df <- cases %>% dplyr::filter(Events>=1) %>% 
  group_by(country, City, study_type) %>%
  summarize(min_month = min(pandemic_month),
            max_month = max(pandemic_month)) %>%
  group_by(country) %>%
  mutate(country_study_number = 1:n(),
         country_total_studies = n())
# see https://stackoverflow.com/questions/36893957/dplyr-generate-row-number-row-position-in-group-by

gap_betw_studies <- 0.15

ggplot() +
  geom_bar(data = df8, stat="identity",
           mapping = aes(x = date_ym, fill = variant, y=nor)) +
  scale_fill_manual(values = c("Alpha" = "deepskyblue2", "Pre_alpha" = "lightsteelblue1", "Beta" = "darkseagreen1",
                               "Delta" = "gold", "Epsilon" = "burlywood", "Eta" = "cadetblue2", "Gamma" = "darkgoldenrod1",
                               "Iota" = "cyan3", "Lambda" = "gray70", "Mu" = "peachpuff1", "Omicron" = "plum2",
                               "Theta" = "tomato")) + 
  theme_light() + 
  theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size = 6),
        axis.text.y = element_text(size = 6)) + 
  xlab("Year-Month") + 
  ylab("Relative Frequency") +
  geom_rect(data = studies_df,
            mapping=aes(xmin=min_month-0.5,
                        xmax=max_month+0.5,
                        ymin= 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + gap_betw_studies*(country_study_number - 1) - 0.005,
                        ymax= 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + gap_betw_studies*(country_study_number - 1) + 0.005)) +
  geom_text(data = studies_df,
            mapping=aes(x = 0.5*(min_month + max_month),
                        y = 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + gap_betw_studies*(country_study_number - 1) + 0.05,
                        label = ifelse(is.na(City),
                                       study_type,
                                       paste(City, study_type))),
            size = 1.5) +
  guides(fill = guide_legend(override.aes = list(labels = c("A", "B")))) +
  facet_wrap(~country, nrow = 30, ncol = 1, scale = "free_x")

ggsave(filename = 'data_out/covid_variant_country_20220525.pdf', width = 6, height = 30, units = "in")
```



