---
title: "Vriant_Analysis"
author: "Fouzia Farooq"
date: "6/3/2022"
output: pdf_document
---


```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(ggplot2)
library(RColorBrewer)
library(readxl)
library(xlsx)
library(patchwork)
library(haven)
```

NOTES: json file was converted to csv in python (deepnote.com), then copied the code into BBtext and ran at the command line. 
Then ran the following code. 
Dates are fractions of a year after "year." We a creating a seperate dataframe into a csv that we will bring in R and use lubridate to read the date properly. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


OPENSOURCE
```{r}
frequency <- read.csv('data/Nextstrain_opensource_frequenices.csv')
covid_opensource <- read.csv('data/nextstrain_ncov_open_global_all-time_metadata_20220517.csv')
pivots <- read.csv('data/Nextstrain_opensource_pivotdates.csv', header = FALSE)
study_cases <- read_xlsx('data/variant_study_case_control_COVID_data_ff_20220524.xlsx') # case-comparison country study dates

# Study files
compiled_ipd_data <- read_dta('data/Compiled_IPD_Data_220620.dta') # updated with June 2022 data
agd_wide <- read_dta('data/Compiled_AGD_WIDE_220408.dta') # Has events, country, but no dates.
agd_long <- read_dta('data/Compiled_AGD_220408.dta') # Has events, country, but no dates.
country_id <- read.csv('data/country_id_ipd.csv') # FF created this file
country_id_agd <- read.csv('data/country_id_agd.csv')
```

```{r}
table(compiled_ipd_data$site_id)
```

```{r}
opensource.df <- covid_opensource %>% dplyr::select(strain, date, country, clade_membership)

opensource.df2 <- left_join(frequency, covid_opensource)

pivot_df <- data.frame(week = 0:(length(pivots$V1)-1), pivot = pivots$V1)
pivot_df <- pivot_df %>% dplyr::mutate(date = as.Date(date_decimal(pivot))) #converting the dates in decimal in nextstrain dataset to readable dates.

df <- left_join(opensource.df2, pivot_df, by="week")
```


```{r}
df2 <- df  %>% dplyr::mutate(variant = case_when(clade_membership %in% 
                                                   c("19A", "19B", "20A", "20B", "20C", "20D", "20F", "20G", "20E (EU1)") ~ "Pre_alpha",
                                                 clade_membership %in% 
                                                   c("20I (Alpha, V1)") ~ "Alpha",
                                                 clade_membership %in%
                                                   c("20H (Beta, V2)") ~ "Beta", 
                                                 clade_membership %in%
                                                   c("21B (Kappa)") ~ "Kappa", 
                                                 clade_membership %in%
                                                   c("21D (Eta)") ~ "Eta",
                                                 clade_membership %in% 
                                                   c("21E (Theta)") ~ "Theta",
                                                 clade_membership %in% 
                                                   c("20J (Gamma, V3)") ~ "Gamma",
                                                 clade_membership %in% 
                                                   c("21A (Delta)", "21I (Delta)", "21J (Delta)") ~ "Delta",
                                                 clade_membership %in% 
                                                   c("21C (Epsilon)") ~ "Epsilon",
                                                 clade_membership %in% 
                                                   c("21F (Iota)") ~ "Iota",
                                                 clade_membership %in% 
                                                   c("21G (Lambda)") ~ "Lambda",
                                                 clade_membership %in% 
                                                   c("21H (Mu)") ~ "Mu",
                                                 clade_membership %in% 
                                                   c("21K (Omicron)", "21L (Omicron)", "21M (Omicron)", "22C (Omicron)") ~ "Omicron"))

table(df2$clade_membership, df2$variant, useNA = "always") # These all have a general variant name now. 

```

NORMALIZE FREQ BY COUNTRY
```{r}
df3 <- df2 %>% group_by(country, week, variant) %>% summarize(freq = sum(frequency))
df4 <- df3 %>% group_by(country, week) %>% 
  dplyr::mutate(nor = (freq/sum(freq))) %>%
  dplyr::filter(country %in% c("Kenya", "Uganda", "Spain", "Hong Kong", "Italy",
                               "South Africa", "USA", "Sweden", "Turkey", "China", "United Kingdom", "Chile", "Canada", 
                                "Mexico", "Colombia", "USA", "India", "Africa", "Kenya", "Pakistan", "Bangladesh")) # List does not include DRC or Burkina Faso b/c opensource data is not available thorugh Nextstrain.  This list doesn't include all the african countries or the WHO countries. 

```




```{r}
df5 <- left_join(df4, pivot_df, by="week")

df6 <- df5 %>% dplyr::mutate(date_ym = format(date, format= "%y-%m"))
df7 <- df6 %>% group_by(country, date_ym) %>% dplyr::mutate(max_date = max(date))
df8 <- df7 %>% dplyr::filter(date == max_date)
sort(colnames(compiled_ipd_data))
check.df <- compiled_ipd_data %>% dplyr::select(site_id, person_id, mat_age, preterm_labor)
```


Joining the case-comparison and RFA dates:
```{r}
study_cases$date <- mdy(study_cases$date)
cases <- study_cases %>% dplyr::mutate(date_ym = format(date,format="%y-%m"))
class(study_cases$date)

variants_cases <- left_join(df8, cases, by=c("country", "date_ym"))
```


************************
STARTING STUDY ANALYSIS:
************************
```{r}
df9 <- left_join(compiled_ipd_data, country_id, by="site_id")
df10 <- df9 %>% dplyr::filter(site_id %in% c(4,5,17,10,15,19,20,24,34, 35, 38)) %>% 
  # Added 20, 24, 34, 37 and updated 5,17,10,19
  dplyr::select(site_id, city_country, country, mat_age, date_onset_or_test, covid_sympdate, covid_pos,
                               edd_date, fetuses, parity, gravidity, weight_pre, diab_prepreg, hyperten_prepreg, matdeath,
                              matdeath_date, covid_hosp, covid_icu, preeclampsia, eclampsia, pre_or_eclampsia, 
                              place_abrupt, preterm_labor, haemorrhage, hpd_any) 
# use date_onset_or_test --> If tested for covid, it is date of first positive test.  
#puerto rico, canada, mexico, madrid spain, hong kong, italy, kenya

check.df <- df9 %>% dplyr::select(site_id, mat_age, preterm_labor)

df10 <- df10 %>% dplyr::mutate(date_ym = format(date_onset_or_test,format="%y-%m"))
unique(df10$country)
unique(compiled_ipd_data$site_id)
```

```{r}
## CREATEING A NEW FILE SO I CAN OVERLAY ON GRAPHS AGAIN TO SEE  IF WE HAVE  MORE THAN 1 VRIANT DATA WITH THIS UPDATED FILE FROM ERIN OR NOT.

min(df10$date_onset_or_test, na.rm=TRUE)
df10.2 <- df10 %>% dplyr::select(site_id, city_country, country, date_onset_or_test, date_ym) 
df10.2 <- df10.2 %>% dplyr::mutate(date_onset_or_test = format(date_onset_or_test,format="%m-%d-%y"))
df10.3 <- df10.2 %>% group_by(country) %>% arrange(country, date_ym)
df10.4 <- df10.3 %>% group_by(country, city_country, date_ym) %>% tally()
df10.4 <- df10.4 %>% rename(Events='n', City = "city_country") %>% mutate(study_type = "RFA",
                                                   Study_type_full = "Risk Factor Analysis")
df10.4 <- df10.4 %>% mutate(date = as.Date(paste0("20", date_ym, "-01")))
df10.5 <- df10.4 %>% mutate(pandemic_month = interval("2019-12-01", date) %/% months(1))

# df10.5$date <- format(df10.5$date, format="%y-%m-%d")
df10.6 <- df10.5 %>% dplyr::select(-date_ym)

# write.csv(df10.5, 'data/variant_newdata_dates_ipd_ff_20220627.csv', row.names = FALSE) # writing this file out so I can overlay on the graphs from nextstrain data
study_cases_2 <- study_cases %>% dplyr::select(-study_period)

df10.7 <- rbind(study_cases_2, df10.6)
df10.8 <- df10.7 %>% arrange(country, study_type, date)
```

New map with overlayed CC study dates and RFA study dates: 06/27/2022 - latest
```{r}

studies_df <- df10.8 %>% dplyr::filter(Events>=1) %>% 
  group_by(country, City, study_type) %>%
  summarize(min_month = min(pandemic_month),
            max_month = max(pandemic_month)) %>%
  group_by(country) %>%
  mutate(country_study_number = 1:n(),
         country_total_studies = n())
# see https://stackoverflow.com/questions/36893957/dplyr-generate-row-number-row-position-in-group-by

gap_betw_studies <- 0.15

ggplot() +
  geom_bar(data = df8, stat="identity",
           mapping = aes(x = date_ym, fill = variant, y=nor)) +
  scale_fill_manual(values = c("Alpha" = "deepskyblue2", "Pre_alpha" = "lightsteelblue1", "Beta" = "darkseagreen1",
                               "Delta" = "gold", "Epsilon" = "burlywood", "Eta" = "cadetblue2", "Gamma" = "darkgoldenrod1",
                               "Iota" = "cyan3", "Lambda" = "gray70", "Mu" = "peachpuff1", "Omicron" = "plum2",
                               "Theta" = "tomato")) + 
  theme_light() + 
  theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size = 6),
        axis.text.y = element_text(size = 6)) + 
  xlab("Year-Month") + 
  ylab("Relative Frequency") +
  geom_rect(data = studies_df,
            mapping=aes(xmin=min_month-0.5,
                        xmax=max_month+0.5,
                        ymin= 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + gap_betw_studies*(country_study_number - 1) - 0.005,
                        ymax= 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + gap_betw_studies*(country_study_number - 1) + 0.005)) +
  geom_text(data = studies_df,
            mapping=aes(x = 0.5*(min_month + max_month),
                        y = 0.5 - 0.5*gap_betw_studies*(country_total_studies-1) + gap_betw_studies*(country_study_number - 1) + 0.05,
                        label = ifelse(is.na(City),
                                       study_type,
                                       paste(City, study_type))),
            size = 1.5) +
  guides(fill = guide_legend(override.aes = list(labels = c("A", "B")))) +
  facet_wrap(~country, nrow = 30, ncol = 1, scale = "free_x")

# ggsave(filename = 'data_out/covid_variant_country_20220627.pdf', width = 6, height = 30, units = "in") # DON'T NEED TO WRITE THIS OUT EVERYTIME.

# write.csv(df8, 'data/nextstrain_frequency_data_written_ff_20220610.csv', row.names = FALSE)
```


Making frequency dataset wider and then we will join it onto compiled_ipd data
```{r}
df8.1 <- df8 %>% group_by(country, date_ym) %>%
  arrange(desc(freq)) %>% 
  mutate(dominant_variant = if_else(first(freq)==0,
                                    "no_strain", 
                                    first(variant))) %>% 
  ungroup() %>%
  arrange(country, date_ym) # first arranges the strains in descending order then pulls out the strain name for the first one b/c now it has the greatest frequency. Then we ungrouped and rearranged so we can see by country and date order. 
  

df8_wide <- df8.1 %>% 
  dplyr::select(-c(freq, week, date, max_date, pivot)) %>% 
  pivot_wider(names_from = variant, values_from = nor, values_fill = 0, names_prefix = "freq_") # pivot wider and adds a prefix "freq" to the variable in "values_from".
# Pivot wider comes from tidyr.
```

Merge the frequency dataset (wide) on our study dataset (IPD dataset)
```{r}
df11 <- left_join(df10, df8_wide, by=c('country', 'date_ym'))
```

First pivot longer. (have a single column for the outcomes)
```{r}
colnames(df12)
df12 <- df11 %>% pivot_longer(cols = c(diab_prepreg, hyperten_prepreg, matdeath, covid_hosp, covid_icu,
                                       preeclampsia, eclampsia, pre_or_eclampsia, place_abrupt, preterm_labor,
                                       haemorrhage, hpd_any),
                              names_to = "outcome", values_to = "events")
# There are lots of "88" in the outcomes --> converting to NA
df12 <- df12 %>% mutate(events = if_else(events==88, as.numeric(NA), events))
df13 <- df12 %>% dplyr::filter(!is.na(events))
```

Groupby and summarize
```{r}
df14 <- df13 %>% group_by(country, dominant_variant, outcome) %>%
  summarize(events = sum(events, na.rm = FALSE),
            tot_n = n(),
            incidence = events/tot_n)

write.xlsx(as.data.frame(df14), file='data_out/variants_ipd_results_20220706.xlsx', sheetName="dominant_strain_country", row.names = FALSE)

df14.1 <- df13 %>% group_by(dominant_variant, outcome) %>%
  summarize(events = sum(events, na.rm = FALSE),
            tot_n = n(),
            incidence = events/tot_n)

write.xlsx(as.data.frame(df14.1), file='data_out/variants_ipd_results_20220706.xlsx', sheetName="dominant_strain_outcome", row.names = FALSE, append=TRUE)
```

```{r}
# This is more useful

df14.1 %>% 
  filter(dominant_variant!="Beta") %>% # beta numbers are very low. 
  ggplot() +
  geom_col(aes(x=dominant_variant, y=incidence, fill=outcome), 
           width = 0.65,
           position="dodge")

df14.1 %>% 
  filter(dominant_variant!="Beta") %>% 
  ggplot() +
  geom_col(aes(x=outcome, y=incidence, fill= dominant_variant), 
           width = 0.7, 
           position="dodge") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```

Making data wider again for regression analysis
```{r}
ipd_wide14.1 <- df14 %>% pivot_wider(names_from = outcome, values_from = events)
```




For modeling, using wide ipd dataset. 
```{r}
summary(mod1 <- glm(preterm_labor ~ variant_mod, family = "poisson", data=ipd_wide14.1))

```


USING *AGGREGATE* DATA: older file. 
```{r}
agd_df <- agd_long %>% dplyr::filter(grepl("calmonth", Strata)) # pulls out the rows that start with "calmonth" in the strata col. 
agd_df <- agd_df %>% separate(Strata, c("calmonth", "calmonth_date"), "_")
agd_df <- agd_df %>% mutate(date_ym = as.Date(paste0(substring(calmonth_date, 1,4), "-", substring(calmonth_date, 5), "-01"),
                                              format="%Y-%m-%d")) # substring pulls out the first 4 characters of the string first. 
agd_df <- agd_df %>% mutate(date_ym = format(date_ym, format = "%y-%m"))

agd_df <- rename(agd_df, site_id = "Site")
```

Merge the frequency dataset (wide) on our study dataset (Aggregate dataset)
```{r}
# country_id_agd <- agd_df %>% dplyr::select(site_id, Site_name) %>% distinct()
# write.csv(country_id_agd, 'data/country_id_agd.csv', row.names = FALSE) # I wrote this out and will update, so don't need this code.

agd_df2 <- left_join(agd_df, country_id_agd, by="site_id")
agd_df2 <- agd_df2 %>% dplyr::select(-Site_name.y)

agd_df3 <- left_join(agd_df2, df8_wide, by=c('country', 'date_ym'))
```

Summarize data
```{r}
#  "Events" column is actually the outcome and the "Total" is the denominator for that Strata
# So for example, if the Outcome is "Preterm Birth" and the Strata is "20203", then "Events" is the number of preterm births among livebirths with COVID-19 onset in March 2020, and "Total" is the total number of livebirths with COVID-19 onset in March 2020

#TODO calmonth_date has "m" that needs to be deleted - these are missing. We don't know what calmonth the outcome happened. 
table(agd_df3$Outcome)

agd_df3.1 <- agd_df3 %>% dplyr::filter(country==c("USA", "Canada", "Mexico", "Spain", "Hong Kong", "Italy", "Kenya"))
agd_df3.1 <- agd_df3.1 %>% dplyr::mutate(variant_mod = if_else(dominant_variant=="Pre_alpha", "Pre_alpha", "other_variant")) 
agd_df3.1$variant_mod <- factor(agd_df3.1$variant_mod, levels=c("other_variant", "Pre_alpha")) # factorized this variable
levels(agd_df3.1$variant_mod)

# wide dataset
agd_wide3.1 <- agd_df3.1 %>% pivot_wider(names_from = Outcome, values_from = Events)

agd_results <- agd_df3.1 %>% group_by(country, dominant_variant, Outcome) %>% summarize(count_condition=sum(Events),
                                                                       tot_n = sum(Total),
                                                                       incidence = count_condition/tot_n) %>% 
  ungroup()
# write.xlsx(as.data.frame(agd_results), file='data_out/variants_agd_results_20220610.xlsx', sheetName="dominant_strain", row.names = FALSE)

agd_results2 <- agd_df3.1 %>% group_by(country, variant_mod, Outcome) %>% summarize(count_condition=sum(Events),
                                                                       tot_n = sum(Total),
                                                                       incidence = count_condition/tot_n) 
# write.xlsx(as.data.frame(agd_results2), file='data_out/variants_agd_results_20220610.xlsx', sheetName="prealpha_vs.other", append=TRUE, row.names = FALSE)

```

```{r}
p1 <- agd_wide3.1 %>% 
  ggplot() +
  geom_histogram(aes(x=ICUadmit1/Total, fill=dominant_variant)) 

p2 <- agd_wide3.2 %>% 
  ggplot() +
  geom_histogram(aes(x=prdeath1/Total, fill=dominant_variant)) # maternal death

p3 <- agd_wide3.2 %>% 
  ggplot() +
  geom_histogram(aes(x=haemorrhage1/Total, fill=dominant_variant))

p4 <- agd_wide3.2 %>% 
  ggplot() +
  geom_histogram(aes(x=extremesga1/Total, fill=dominant_variant))

p5 <- agd_wide3.2 %>% 
  ggplot() +
  geom_histogram(aes(x=preterm1/Total, fill=dominant_variant))

p6 <- agd_wide3.2 %>% 
  ggplot() +
  geom_histogram(aes(x=c_intrap1/Total, fill=dominant_variant))

p1/p2/p3/p4/p5/p6


ggsave('data_out/variant_rates_20220610.pdf', plot=last_plot(), height = 13, width=4)
```

For modeling, using wide aggregated dataset. 
```{r}
summary(mod1 <- glm(preterm1 ~ variant_mod + country, family = "poisson", data=agd_wide3.2))

```


